E Coli Project Paper Incredibly Rough Draft

### 2013-11-09 paper/paper.tex
a few edits of paper Tested everything on collins.

\newcommand{\fixme}[1]{\red{[#1]}}

@@

\section{Introduction}
It is vital that during the process of bacterial cell division a cell
avoid minicelling, or splitting into daughter cells with lopsided
volumes.  During this process a long FtsZ polymer chain develops on
the cell wall in the center region of the cell that dictates the plane
of division\cite{adams2009bacterial}
\cite{lutkenhaus2007assembly}. Previous experimental studies have
shown that the MinC protein, known to inhibit the FtZ
polymer\cite{shen2010examination}, exhibits a pole to pole oscillatory
behavoir in conjunction with the MinD and MinE proteins, while the
oscillating MinE tends to be more localized in the center of the cell
\cite{hu1999topological} \cite{fu2001mine} \cite{shapiro2009and}
\cite{yu1999ftsz} \cite{raskin1999rapid}The MinC protein will then have a higher time
averaged concentration in the cell poles as opposed to the center
region of the cell, aiding in prohibiting the FtZ from developing in the wrong
region.

A significant amount of work has been done to develop protein reaction
and diffusion models that exhibit accurate macroscopic dynamics of the
MinD protein system. Early models involved free proteins that affect
eachothers' rates of diffusion and membrane attachement but do not
combine into compound states\cite{meinhardt2001pattern} .  In 2003
Huang \cite{huang2003dynamic} improved upon this work with a simple
model based on MinD-MinE combination, ATPase hydrolysis, and MinD
membrane attechent that when simulated exhibits accurate MinD
oscillations in cylindrical cells\cite{huang2003dynamic}. In this
model cytoplasmic MinD is more likely to attach to the membrane when
MinD is already clustered there (following observed non-linear
attachement of minD on the cell membrane), and is stationary once
attached.  A number of studies have used an approach similar in that
they do not rely on the ability of MinD to move along the walls and
cluster ~\cite{kruse2007experimentalist} \cite{meinhardt2001pattern}
\cite{drew2005polymerization} \cite{fange2006noise}
\cite{kerr2006division}, while studies have been made as well of
models which rely on MinD mobility and attraction on the cell
membrane\cite{kruse2002dynamic}\cite{howard2005cellular}.  Variations
of the Huang 2003 model that stochasticly account for variations of
molecular interaction \cite{fange2006noise} and as well monte-carlo
simulations that implement stochastic version of Huang's mean filed
reaction rates confirm the major results obtained by Huang's model and
more successfully predict experimentally observed oscillations in
round cell phenotypes.\cite{drew2005polymerization}
\cite{fange2006noise} \cite{huang2004min} Biochemical models of
broader scope have also been used to study the MinD system and show
consistent results.\cite{arjunan2010new}.  However, in general the
results of the stochastic and monte-carlo simulations are similar to
those given by Huang's mean field results.

@@

It is vital for the mechanism by which the cell accurately divides to
be robust and indeed it has been shown to
be\cite{touhami2006temperature}.  Studies have shown that the
oscillations can effectively find poles in abberent cell
shapes~\cite{corbin2002exploring} \cite{juarez2010changes}.  Varma et
al. have studied three pronged connected tube shapes, both
experimentally and in simulation\cite{varma2008min} and have shown
that even in these oscillations develop in regular patterns. However,
Mannik et al. have shown that there are limits to this
robustness. Forcing cells into flattened, irregular cell shapes
adversly effects the Min system's ability to maintain their regular
oscillatory behavior \fixme{this link doesn't work = cite{mannik2012robustness}}
\cite{mannik2010bacteria} \cite{mannik2009bacterial}.  We follow the
experimental work of Mannik et and use Huang's mean field differential
equation reaction model to explore the model's ability to find regular
oscillatory behavoir in a series of abberent shapes.

\subsection{What is the MinD system and why is it important?}
\subsection{How proteins move in cell}
\section{Methods and Initial Conditions}
The model for the behavior of the MinD and MinE proteins inside the
cell implements the same set of 5 reaction-diffusion equations studied in
the work of Huang et al (equations 1, 2, 3, 4, and
5)\cite{huang2003dynamic}. A 3d grid is constructed in cartesian
coordinates with a grid spacing of .05 $\mu$m. We define our cell
shapes and solve the reaction-diffusion equations numerically to
observe the time evolution of the MinD and MinE concentrations inside
the cell.

We start our cells with the MinD and MinE concentrations that are
reported as wild type concentrations by Huang et al . We as well use the same reaction
diffusion constants and and reaction rates,
\begin{gather*} %format better
  \mathcal{D}_D = \mathcal{D}_{E}  = 2.5 \mu \textrm{m$^2$ / sec}, \\
  \sigma_D^{\textrm{ADP $\rightarrow$ ATP}}  = 1/\textrm{sec},  \sigma_D = 0.025 \mu \textrm{m/sec}, \\
  \sigma_{dD}  = 0.0015 \mu \textrm{m$^3$/sec}, \\
  \sigma_{de}  = 0.7/\textrm{sec}, \sigma_E = 0.093 \mu \textrm{m$^3$/sec}.
\end{gather*}

We study cell shapes that are similar to those studied by Huang and
also new, flattened pancake-like shapes, inspired by the experimental work of
Mannik \cite{mannick2012robustness}.  Our pill shapes differ from
Huangs in that they are cylinders with hemisphere endcaps instead of
pure cylindridrical shapes.  Our cylinderidrical radius is $0.50\mu$
and the lengths of our cells (measured between the tips of the
endcaps) are $5\mu$, $4\mu$, $3\mu$, and $2.5\mu$.

With our flattened cell shapes we mean to study cells that are
parrallel to those lodged into crevices by Mannik so we give them
height of $0.25\mu$ \cite{mannick2012robustness}.  Viewed from the top
down the cells will have the shapes described below and viewed from
the side they have at their edges a semicircular protrusion (one may
imagine the edges of a pancake).

@@

Kubitschek\cite{kubitschek1990cell} \cite{kubitschek1968linear} has
shown in multiple experiments that at the time of cell division cells
have a volume that is within a range of roughly $1\mu m^3$ to $2\mu
m^3$.  We follow Huang's simulations\cite{huang2003dynamic} and
Mannik's experiments \fixme{this link doesn't work = cite{mannik2012robustness}} and model cells that
are slightly larger than this range.  Amoung our flattened cells the
two dimensional length scales are tuned so that every total cell
volume is very close to $3\mu^3$.  \fixme{should we just leave it to
  the reader to figure out for themselves what our pill shape volumes
  are?}

To interpret the results, we generated several different plot views of
the printed simulation data. These plots included a time averaged view
of the protein densities in the cell; a plot tracking the location of
protein concentrations that were global maxima in space and local maxima in
time; and an animated view that showed the actual dispersion of
protein concentrations in the cell over time.
\subsection{Mathematical Model}
\section{Specific Results}
\subsection{Pill Shape}

Our simplest shape is the naturally occuring pill shape.  We peice this shape
together as two hemispherical endcaps attached on either end of a
cylinder.  This shape follows the early simulations of Huang et al but
differs in that we have added the end caps for a more natural shape,
expecting similar results.  We test cells of radius $.5\mu$ and of
lengths $2\mu$, $3\mu$, and $5\mu$ between both tips of the end
caps. In all these models we observe quick establishement (within a
few periods) of extremely regular oscillatory behavoir that lasts for
any length of time, with periods that closely agree with the results
of Huang et al.  The following table shows the periods of oscillations
at different cell lengths.

This simple model is a good starting point for observing in detail the
dynamic interaction between the different stages of protein that lead
to their qualitatively distinct behavoir. Figure \ref{image-p} shows
the process in a series of frames taken from simulation animations.
Each frame is 2.5 seconds ahead of the last.  The causal chain is
cyclical so we can begin our observations at any stage in the cycle.
We consider first MinD:ATP floating freely in the cytoplasm.  The
protein will not react with any other floating protein, and it will
not spontenously shift back into the MinD:ADP stage of the cycle.  It
will diffuse until it attaches to a cell membrane.  When compared to a
protein that is in the middle cylindrical section of the cell, a
protein that is in an end cap section is more likely to bump into a
wall in a given number of diffusive steps, since there is a greater
cross section of possible movement that is covered by a nearby wall.
All else being equal, proteins will collect with a higher rate on the
caps.  This effect is made more pronounced by the MinD:ATP tendency to
attract and cluster.  When adjacent to the membrane, the rate of
binding to the membrane is porportional to the amount of MinD already
bound there.  Initially, then, there is a strong tendency of MinD to
cluster in the polar regions of the cell.

The next observed mechanism is the formation of MinE 'rings' on the
membranes of the polar caps.  As the freely
diffusing MinE enters the cap portion it begins to encounter
concentrated membrane bound MinD.  As it bumps into the walls here it
binds to this MinD and remians for a time on the walls in compound
form.  When it is released, it can diffuse naturally outward or inward
towards the cell.  If it diffuses inward, there is less cross
sectional wall for it to bump into and also less concentrated MinD:ATP
to react with.  If it diffuses outward, the curvature of the cell wall
cuases there to be a greater chance that it will bump into the cell
membrane (for the same reason outlined for the MinD:ATP above) and if
there is MinD:ATP bound there, there is a chance that it will bind to
that protein and react again.  If it does diffuse in this directin, it
will likely react with MinD:ATP before getting very far into the
encap.  Thus after dettaching from the membrane at the edge of the
MinD:ATP encap cluster, the MinE is most likely to travel a small
distance further into the polar region and then attach and react again
with MinD:ATP protein there.  This process results in the observed
qualitative 'encroaching MinE ring' behavoir that releases MinD:ADP
back into the cytoplasm.  This MinD:ADP will diffuse for a time before
spontanosly uindergowing hydrolysis \fixme{is this right?} and
reaching the MinD:ATP state that can bind to the membrane again.
There is a strong chance that it has by this time diffused out of the
polar region and is on its way to the opposite polar region of the
cell.  If it hasn't then it will reattach to the MinD:ATP on the cell
wall deeper into the endcap and will eventually be converged upon by
the MinE ring, only to be released again in MinD:ADP form.

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-400-50-0-0-1500.pdf}
  \caption{Frames from an animation showing a series of contour plots
    of the concentration of proteins in different stages over time.
    The density of proteins are integrated along the axis normal to
    the page. The frames occur 2.5 seconds apart. \fixme{this looks
      assymetric! That's a big problem.}}
  \label{image-p}
\end{figure*}

Consider Figure \ref{image-p}.  At 0 seconds we can see cytoplasmic
nATP roughly evenly distributed throughout the cell.  There is nADP
concentrated in the left pole and it will transform into nATP.  At 2.5
seconds we see membrane bound MinD:ATP form on the right side of the
cell.  It avoids the left side because there is a concentration of
cytoplasmic MinE on the left side and compared to the right side,
which is empty of MinE, MinD:ATP attaching to the wall on the left has
no time to collect and cluster before reacting.  One can see how as
the MinE spreads out rightward, the the clustered membrane bound
MinD:ATP stays just at the edge of it, always to the right.  Where the
two meet there is at every time step a spike in membrane bound
MinD-MinE:ATP, resulting from the MinE reaction with the MinD on the
wall.  One can see from roughly 5 seconds to 17.5 seconds the
cytoplasmic MinE chases in a way the MinD towards the right, reacting
with it as it clusters on the wall.  Upon reaching the right side, the
MinD is in a way trapped, and the MinE now concentrated in that pole
will react until the MinD has diffused back towards the other side and
there is no more membrane bound MinD:ATP to react with.  The cycle
then begins again.

\begin{figure}
  %%\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-400-50-0-0-1500}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\mu m$ by 1$\mu m$ pill shaped cell (above) and a
    3$\mu m$ by 1$\mu m$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{box-p}
\end{figure}

@@

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period According to Pill Length}
\end{table}

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/ave-plot_D--p-200-50-0-0-1500}
  \caption{Average protein density per unit area of wall for the wall
    attached stages of the protein cycle.  Shown in the right, middle,
    and left parts of a 5$\mu m$ by 1$\mu m$ pill shaped cell (above)
    and a 3$\mu m$ by 1$\mu m$ pill shaped cell (below).}
  \label{ave-per-area-plot-pill}
\end{figure}

Figure \ref{ave-per-area-plot-pill} shows the ave protein per unit
area that is wall attached in the middle, right, and left sections of
the pill shaped cell.  This is an important test of the central idea
behind the utility of the Min protein system.  The system is meant to
place Min proteins on the walls in heavier concentrations in the caps
of the cell and lower concentrations in the middle section, so that
the FtZ polymer may build up along the center walls.  We can see from
the plot of the 3$\mu m$ by 1$\mu m$ cell that the MinD proteins reach
concentration peaks in the caps that are higher than peaks in the
center.  \fixme{is the next part true?} Integrals of these plotted values show that the time averaged
concentrations are likewise higher in the caps than in the center.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-601-601-601-1500}
  \caption{Total protein fluctuation of an equilateral triangle with
    larger dimensions (each side is roughly $6\mu m$.  The vertical
    axis shows stacked the total number of proteins that are of four
    different compound stages and in the different sections of the
    cell. The time limits have been set to show the broken down
    section of the simulation.}
  \label{box-triangle}

@@

Simulations in which the two planar dimensions are 2 and then 3 times
the size (cells roughly $6\mu m$ and $9\mu m$, respectively, from top
to bottom) show no stable regular periodic pattern \fixme{although
  they haven't really run long enough to exhibit much.  They certainly
  won't show the double oscillation pattern, however.}  The animations
of $6\mu m$ cells show that the MinD acts as a sort of traveling
cluster that is constantly being eaten away be MinE behind it.  When
this cluster enters a region of high curvature that ends in a dead end
corner it performs the polar maxima behavoir seen in the other
simulations, regrouping out away from the corner where it travels in
the opposite direction while being ``chased'' away by the MinE that
has itself maximized in the corner.  If it has just maximized in the
top corner, it regroups traveling downward and then naturally splits
into two groups that travel out and form maxima in the two opposite
bottom corners.  After maximizing in these two corners the two groups
recombine again in the center and travel in some direction together as
one cluster.  Simulations of the $9\mu m$ cells show a similar roaming
cluster.  Although here, the maxima occur on non-polar walls often,
and will often split into two roaming clusters that interact less
quickly and are less correlated than those seen in the $6\mu m$ cell.

>>>>

\newcommand{\fixme}[1]{\red{[#1]}}
\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}

\newcommand\micron{\ensuremath{\mu\text{m}}}

@@

\section{Introduction}
It is vital that during the process of bacterial cell division a cell
avoid minicelling, or splitting into daughter cells with lopsided
volumes.  During this process a long FtsZ polymer chain develops on
the cell wall in the center region of the cell that dictates the plane
of division~\cite{adams2009bacterial,
  lutkenhaus2007assembly}. Previous experimental studies have shown
that the MinC protein, known to inhibit the FtZ
polymer\cite{shen2010examination}, exhibits a pole to pole oscillatory
behavoir in conjunction with the MinD and MinE proteins, while the
oscillating MinE tends to be more localized in the center of the
cell~\cite{hu1999topological, fu2001mine, shapiro2009and, yu1999ftsz,
  raskin1999rapid}. The MinC protein will then have a higher time
averaged concentration in the cell poles as opposed to the center
region of the cell, aiding in prohibiting the FtZ from developing in
the wrong region.

It is vital for the mechanism by which the cell accurately divides to
be robust and indeed it has been shown to
be\cite{touhami2006temperature}.  Studies have shown that the
oscillations can effectively find poles in abberent cell
shapes~\cite{corbin2002exploring} \cite{juarez2010changes}.  Varma et
al. have studied three pronged connected tube shapes, both
experimentally and in simulation\cite{varma2008min} and have shown
that even in these oscillations develop in regular patterns. However,
Mannik et al. have shown that there are limits to this
robustness. Forcing cells into flattened, irregular cell shapes
adversly effects the Min system's ability to maintain their regular
oscillatory behavior \fixme{this link doesn't work =
  cite{mannik2012robustness}} \cite{mannik2010bacteria}
\cite{mannik2009bacterial}.  We follow the experimental work of Mannik
et and use Huang's mean field differential equation reaction model to
explore the model's ability to find regular oscillatory behavoir in a
series of abberent shapes.  \davidsays{This paragraph could use some
  work, and should reference Mannik early, and emphasize his research.
  First paragraph sets the stage for why our work is important by
  introducing the MinD system.  Second paragraph should explain why we
  are studying the Min system in funny cell shapes.}

A significant amount of work has been done to develop protein reaction
and diffusion models that exhibit accurate macroscopic dynamics of the
MinD protein system. Early models involved free proteins that affect
eachothers' rates of diffusion and membrane attachement but do not
combine into compound states~\cite{meinhardt2001pattern}.  In 2003
Huang improved upon this work with a simple model based on MinD-MinE
combination, ATPase hydrolysis, and MinD membrane attechent that when
simulated exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is more
likely to attach to the membrane when MinD is already clustered there
(following observed non-linear attachement of minD on the cell
membrane), and is stationary once attached.  A number of studies have
used an approach similar in that they do not rely on the ability of
MinD to move along the walls and
cluster~\cite{kruse2007experimentalist, meinhardt2001pattern,
  drew2005polymerization, fange2006noise, kerr2006division}, while
studies have been made as well of models which rely on MinD mobility
and attraction on the cell membrane~\cite{kruse2002dynamic,
  howard2005cellular}.  Variations of the Huang 2003 model that
stochasticly account for variations of molecular interaction
\cite{fange2006noise} and as well monte-carlo simulations that
implement stochastic version of Huang's mean filed reaction rates
confirm the major results obtained by Huang's model and more
successfully predict experimentally observed oscillations in round
cell phenotypes~\cite{drew2005polymerization, fange2006noise,
  huang2004min}. Biochemical models of broader scope have also been
used to study the MinD system and show consistent
results.\cite{arjunan2010new}.  However, in general the results of the
stochastic and monte-carlo simulations are similar to those given by
Huang's mean field results.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{reactions}
  \caption{Reactions included in the model of Huang \emph{et
      al.}~\cite{huang2003dynamic}.}\label{fig:reactions}
\end{figure}

\section{Model and Methods}
We implement the model of Huang \emph{et al.}  for the behavior of the
MinD and MinE proteins inside the cell~\cite{huang2003dynamic}.  This
model is defined by a set of five reaction-diffusion equations:
\begin{align}
  \frac{\partial n_{?}}{\partial t} &= \text{\davidsays{Actual differential equations here.}}
\end{align}
These equations define the rates of diffusion and reaction shown in
Figure~\ref{fig:reactions}.  \davidsays{Add more text here, describing
  the system and the reactants.  In particular, we want to state that
  MinD:ADP binds ATP (or is hydrolyzed?... check!) to form MinD:ATP,
  which then gets stuck to the cell membrane, which is more likely
  when there is already MinD:ATP bound to the membrane.  The MinE then
  binds to the membrane-bound MinD:ATP.  Finally the whole complex can
  fall off the membrane, at which stage you end up with MinD:ADP and
  are ready to start over.  This is also where we define our
  terminology for the volume and surface densities.}

The reaction shown in Figure~\ref{fig:reactions} is cyclical so we can
begin our observations at any stage in the cycle (which is presented
for the natural cell shape in Fig.~\ref{image-p} below).  We consider
first MinD:ATP floating freely in the cytoplasm.  The protein will not
react with any other floating protein, and it will not spontenously
shift back into the MinD:ADP stage of the cycle.  It will diffuse
until it attaches to a cell membrane.  This is will happen most
rapidly in regions of high curvature, where diffusion is more likely
to lead to an encounter with a wall.  This effect is made more
pronounced by the MinD:ATP tendency to attract and cluster on the
membrane, as reflected in Equation~\ref{eq:d-on-wall}, which shows
that when adjacent to the membrane, the rate of binding to the
membrane is porportional to the amount of MinD already bound to the
wall.  Initially, then, there is a strong tendency of MinD to cluster
on the polar regions of the cell, where the curvature is highest.

We next see the formation of MinE `rings' on the membranes of the
polar caps.  As the freely diffusing MinE enters the cap portion it
begins to encounter concentrated membrane bound MinD.  The MinE binds
to this MinD and remians for a time on the walls in compound form.
When it is released, it can diffuse naturally outward or inward
towards the cell.  If it diffuses inward, there is less cross
sectional wall for it to bump into and also less concentrated MinD:ATP
to react with.  If it diffuses outward, the curvature of the cell wall
cuases there to be a greater chance that it will encounter the cell
membrane (for the same reason outlined for the MinD:ATP above) and if
there is MinD:ATP bound there, there is a chance that it will bind to
that protein and react again.  If it does diffuse in this directin, it
will likely react with MinD:ATP before getting very far into the
encap.  Thus after dettaching from the membrane at the edge of the
MinD:ATP encap cluster, the MinE is most likely to travel a small
distance further into the polar region and then attach and react again
with MinD:ATP protein there.  This process results in the observed
qualitative 'encroaching MinE ring' behavoir that releases MinD:ADP
back into the cytoplasm.  This MinD:ADP will diffuse for a time before
spontanosly uindergowing hydrolysis \fixme{is this right?} and
reaching the MinD:ATP state that can bind to the membrane again.
There is a strong chance that it has by this time diffused out of the
polar region and is on its way to the opposite polar region of the
cell.  If it hasn't then it will reattach to the MinD:ATP on the cell
wall deeper into the endcap and will eventually be converged upon by
the MinE ring, only to be released again in MinD:ADP form.

 A 3d grid is constructed in cartesian
coordinates with a grid spacing of .05\micron. We define our cell
shapes and solve the reaction-diffusion equations numerically to
observe the time evolution of the MinD and MinE concentrations inside
the cell.

We start our cells with the MinD and MinE concentrations that are
reported as wild type concentrations by Huang et al . We as well use the same reaction
diffusion constants and and reaction rates,
\begin{gather*} %format better
  \mathcal{D}_D = \mathcal{D}_{E}  = 2.5 \micron^2/\text{sec}, \\
  \sigma_D^{\textrm{ADP $\rightarrow$ ATP}}  = 1/\textrm{sec},  \sigma_D = 0.025 \micron /\textrm{sec}, \\
  \sigma_{dD}  = 0.0015 \micron^3/ \textrm{sec}, \\
  \sigma_{de}  = 0.7/\textrm{sec}, \sigma_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}

\section{Cell shapes}

In order to address the limits at which the stability of the
oscillations of the Min system break down, as observed by Mannik
\emph{et al.}~\cite{mannik2012robustness}, we model the Min system in
a variety of cell shapes.  We begin with natural pill-shaped cells,
and the proceed to examine a number of flattened pancake-like shapes,
which reflect the experiment of Mannik \emph{et al.}, in which
bacteria were confined within a thin slit~\cite{mannik2012robustness}.
Our pill shapes differ from those of Huang \emph{et al.} in that they
are cylinders with hemisphere endcaps instead of pure cylindridrical
shapes. \davidsays{Double-check the Huang paper on their shapes.}  Our
cylinderidrical radius is $0.50\micron$ and the lengths of our cells
(measured between the tips of the endcaps) are $5\micron$, $4\micron$, $3\micron$,
and $2.5\micron$.

With our flattened cell shapes we mean to study cells that are
parrallel to those lodged into crevices by Mannik so we give them
height of $0.25\micron$ \cite{mannik2012robustness}.  Viewed from the top
down the cells will have the shapes described below and viewed from
the side they have at their edges a semicircular protrusion (one may
imagine the edges of a pancake).

@@

Kubitschek has shown in multiple experiments that at the time of cell
division cells have a volume that is within a range of roughly
$1\micron^3$ to $2\micron^3$~\cite{kubitschek1990cell,
  kubitschek1968linear}.  We follow Huang's
simulations\cite{huang2003dynamic} and Mannik's experiments
\fixme{this link doesn't work = cite{mannik2012robustness}} and model
cells that are slightly larger than this range.  Amoung our flattened
cells the two dimensional length scales are tuned so that every total
cell volume is very close to $3\micron^3$.  \fixme{should we just
  leave it to the reader to figure out for themselves what our pill
  shape volumes are?}

To interpret the results, we generated several different plot views of
the printed simulation data. These plots included a time averaged view
of the protein densities in the cell; a plot tracking the location of
protein concentrations that were global maxima in space and local maxima in
time; and an animated view that showed the actual dispersion of
protein concentrations in the cell over time.
\section{Specific Results}

\subsection{Pill Shape}

\begin{figure*}
  \includegraphics[width=\textwidth]{../data/shape-p/plots/image-plot--p-400-50-0-0-1500.pdf}
  \caption{Frames from an animation showing a series of contour plots
    of the concentration of proteins in different stages over time.
    The density of proteins are integrated along the axis normal to
    the page. The frames occur 2.5 seconds apart. \fixme{this looks
      assymetric! That's a big problem.} \davidsays{I like this plot,
      but would rather see it improved in a few ways.  I noticed the
      implementation combines together bitmaps.  Nicer would be to use
      subplots within matplotlib, so we output the plot directly as a
      pdf.  Then we could also arrange things so there is much less
      vertical white space.  It'd be nicer to have labels at the left
      and a time axis at the bottom, with no labels by each image.  I
      know this will take a while, so it's not yet a high priority.}}
  \label{image-p}
\end{figure*}

We begin with the naturally occuring pill shape.  We peice this shape
together as two hemispherical endcaps attached on either end of a
cylinder.  This shape follows the early simulations of Huang \emph{et
  al.} but differs in that we have added the end caps for a more
natural shape, expecting similar results.  We test cells of radius
$.5\micron$ and of lengths $2\micron$, $3\micron$, and $5\micron$
between both tips of the end caps. In all these models we observe
quick establishement (within a few periods) of extremely regular
oscillatory behavoir that lasts indefinitely, with periods that
closely agree with the results of Huang \emph{et al.}
Table~\ref{tab:pill-periods} shows the periods of oscillations
at different cell lengths.

\begin{table}
  \begin{tabular}{|r|c|c|c|l|}
    \hline
    Length($\mu$) & 2.50 & 3.00 & 4.00 & 5.00\\
    \hline
    Period(sec) & sim & 33 & 38 & 48 \\ \cline{2-2}
    \hline
  \end{tabular}
  \caption{Period According to Pill Length}\label{tab:pill-period}
\end{table}

This simple model is a good starting point for observing in detail the
dynamic interaction between the different stages of protein that lead
to their qualitatively distinct behavoir. Figure \ref{image-p} shows
the process in a series of frames taken from simulation animations.
Each frame is 2.5 seconds ahead of the last.

Consider Figure \ref{image-p}.  In the beginning, the cytoplasmic
nATP is roughly evenly distributed throughout the cell.  There is nADP
concentrated in the left pole and it will transform into nATP.  At 2.5
seconds we see membrane bound MinD:ATP form on the right side of the
cell.  It avoids the left side because there is a concentration of
cytoplasmic MinE on the left side and compared to the right side,
which is empty of MinE, MinD:ATP attaching to the wall on the left has
no time to collect and cluster before reacting.  One can see how as
the MinE spreads out rightward, the the clustered membrane bound
MinD:ATP stays just at the edge of it, always to the right.  Where the
two meet there is at every time step a spike in membrane bound
MinD-MinE:ATP, resulting from the MinE reaction with the MinD on the
wall.  One can see from roughly 5 seconds to 17.5 seconds the
cytoplasmic MinE chases in a way the MinD towards the right, reacting
with it as it clusters on the wall.  Upon reaching the right side, the
MinD is in a way trapped, and the MinE now concentrated in that pole
will react until the MinD has diffused back towards the other side and
there is no more membrane bound MinD:ATP to react with.  The cycle
then begins again.

\begin{figure}
  %%\includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-400-50-0-0-1500}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/box-plot_D--p-200-50-0-0-1500}
  \caption{Total protein fluctuation in the right, middle, and left
    parts of a 5$\micron$ by 1$\micron$ pill shaped cell (above) and a
    3$\micron$ by 1$\micron$ pill shaped cell (below).  The vertical axis shows
    stacked the total number of proteins that are of four different
    compound stages and in the right section of the cell (bottom four
    colors), in the middle of the cell (middle four colors) and in the
    left section of the cell (top four colors). The different forms of
    protein changes while the total number of MinD protein in the cell
    remains constant, shown by the fact that the very top line is
    horizontal.}
  \label{box-p}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-p/plots/ave-plot_D--p-200-50-0-0-1500}
  \caption{Average protein density per unit area of wall for the wall
    attached stages of the protein cycle.  Shown in the right, middle,
    and left parts of a 5$\micron$ by 1$\micron$ pill shaped cell (above)
    and a 3$\micron$ by 1$\micron$ pill shaped cell (below).}
  \label{ave-per-area-plot-pill}
\end{figure}

Figure \ref{ave-per-area-plot-pill} shows the ave protein per unit
area that is wall attached in the middle, right, and left sections of
the pill shaped cell.  This is an important test of the central idea
behind the utility of the Min protein system.  The system is meant to
place Min proteins on the walls in heavier concentrations in the caps
of the cell and lower concentrations in the middle section, so that
the FtZ polymer may build up along the center walls.  We can see from
the plot of the 3$\micron$ by 1$\micron$ cell that the MinD proteins reach
concentration peaks in the caps that are higher than peaks in the
center.  \fixme{is the next part true?} Integrals of these plotted values show that the time averaged
concentrations are likewise higher in the caps than in the center.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{../data/shape-triangle/plots/box-plot_D--triangle-25-601-601-601-1500}
  \caption{Total protein fluctuation of an equilateral triangle with
    larger dimensions (each side is roughly $6\micron$.  The vertical
    axis shows stacked the total number of proteins that are of four
    different compound stages and in the different sections of the
    cell. The time limits have been set to show the broken down
    section of the simulation.}
  \label{box-triangle}

@@

Simulations in which the two planar dimensions are 2 and then 3 times
the size (cells roughly $6\micron$ and $9\micron$, respectively, from top
to bottom) show no stable regular periodic pattern \fixme{although
  they haven't really run long enough to exhibit much.  They certainly
  won't show the double oscillation pattern, however.}  The animations
of $6\micron$ cells show that the MinD acts as a sort of traveling
cluster that is constantly being eaten away be MinE behind it.  When
this cluster enters a region of high curvature that ends in a dead end
corner it performs the polar maxima behavoir seen in the other
simulations, regrouping out away from the corner where it travels in
the opposite direction while being ``chased'' away by the MinE that
has itself maximized in the corner.  If it has just maximized in the
top corner, it regroups traveling downward and then naturally splits
into two groups that travel out and form maxima in the two opposite
bottom corners.  After maximizing in these two corners the two groups
recombine again in the center and travel in some direction together as
one cluster.  Simulations of the $9\micron$ cells show a similar roaming
cluster.  Although here, the maxima occur on non-polar walls often,
and will often split into two roaming clusters that interact less
quickly and are less correlated than those seen in the $6\micron$ cell.
